---
title: "p8105 homework1"
author: "Xieyu Liu"
date: "2026-09-19"
output: github_document
---
# Problem1

```{r}
data("penguins", package = "palmerpenguins")
```
The `penguins` dataset contains information on penguins, including species, island, bill length, bill depth, flipper length, body mass, sex, and year.

The dataset has `r nrow(penguins)` rows and `r ncol(penguins)` columns.

The mean flipper length is `r mean(penguins$flipper_length_mm, na.rm=T)` mm.

```{r}
library(ggplot2)
penguin_plot <- ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm,  color = species)) + geom_point()
penguin_plot
ggsave("penguins_scatterplot.png", plot = penguin_plot)
```

# Problem2

```{r}
library(tidyverse)
df = tibble(
  vec_numeric = rnorm(10),
  vec_logical = vec_numeric > 0,
  vec_char = c("a","b","c","d","e","f","g","h","i","j"),
  vec_factor = factor(c("A","B","C","C","B","A","A","B","C","C"))
)
mean(pull(df, vec_numeric))
mean(pull(df, vec_logical))
mean(pull(df, vec_char))
mean(pull(df, vec_factor))
```
Numeric and logical vectors can be used to calculate the mean, while character and factor vectors cannot.

```{r eval=FALSE}
as.numeric(df$vec_logical)
as.numeric(df$vec_char)
as.numeric(df$vec_factor)
```
Logical and factor vectors can be converted to numeric vectors, while character vectors cannot. R treats `TRUE` as 1 and `FALSE` as 0, which explains why the mean of the logical vector can be calculated. A factor can be converted to numeric values, but `as.numeric()` returns its internal level codes rather than meaningful numeric values.